A linear inside-outside algorithm for correcting sequencing errors in structured RNA sequences

### 2012-10-07 Recomb/conclusion_RECOMB.tex
Typos Concordance des temps dans la partie results

In this article we presented a new and efficient way of
exploring the mutational landscape of an RNA, benefiting from the
 conserved consensus secondary structures  information,
to identify and fix sequencing errors. We also showed a new
 pseudo energy model, taking into account the nearest-neighbour energy model 
and, to quantify geometrical differences,  the isostericity of all base pairs.

By combining those two approaches,  the 
mutational landscape exploration and the pseudo energy model,
 into \texttt{RNApyro}, we show that we can
correct randomly mutated 5s rRNA sequences. 
As presented in Sec.~\ref{sec:results},
we observe that the models
with larger weights on the
isostericity seems to hold a higher accuracy on the estimation of errors.
Importantly, the implementation is fast enough for practical applications.

We must recall that our approach is restricted to
 the correction of point-wise error in structured regions (i.e. base paired nucleotides).
 Nonetheless it supplements well existing tools by using previously discarded
information holding, as shown, a strong signal.

>>>>

In this article we presented a new and efficient way of
exploring the mutational landscape of an RNA, benefiting from the
 conserved consensus secondary structures  information,
to identify and fix sequencing errors. We introduced a new
 pseudo energy model, both taking into account the nearest-neighbour energy model 
and, to account for geometrical discrepancies,  the isostericity of all base pairs.

By combining these two approaches,  the 
mutational landscape exploration and the pseudo energy model,
 into \texttt{RNApyro}, we obtained a tool to predict the position and 
correct randomly mutated 5s rRNA sequences. 
As presented in Sec.~\ref{sec:results},
we observe that the models
with larger weights on the
isostericity seems to hold a higher accuracy on the estimation of errors.
Importantly, the implementation is fast enough for practical applications.

We must recall that our approach is restricted to
 the correction of point-wise error in structured regions (i.e. base paired nucleotides).
 Nonetheless it should supplement well existing tools, by using previously discarded
information holding, as shown, a strong signal.

### 2012-10-07 Recomb/results_RECOMB.tex
fix typos
Typos Concordance des temps dans la partie results

\begin{table}
\begin{center}
\begin{tabular}{lccc}
$|s|$&\multicolumn{3}{c}{Number of mutations}\\\cline{2-4}
		 			  & 6   &  12  & 24\\\cline{2-4}
100  				& 35s  & 238s & 1023s\\
300  			& 135s & 594s &2460s\\\cline{2-4}
		 						& 25   &    &	50		\\\cline{2-4}
500         & 5400s&       &  21003s    \\
\end{tabular}
\end{center}
\caption{Time to compute all probabilities. The first column indicates the length and  the column indexes indicate the number
 of mutations. $\alpha$ is
set at $0.5$,  $\beta$ to $1.5$ and $|\Omega|=44$.}
\label{tab:time}
\end{table}

@@

We built our data set from the 5S rRNA multiple sequence alignment (MSA) available in the Rfam Database 11.0 (Rfam id: \texttt{RF00001}).
Since our software does not currently implement gaps (mainly because scoring indels is a challenging issue that cannot be fully addressed
in this work),  we clustered together the sequences with identical gap locations. From the $54$ MSAs without gap produced, we selected the
biggest MSA  which contains $130$ sequences (out of $712$ in the original Rfam MSA). Then, in order to avoid any bias, we used \texttt{cd-hit}
\cite{Li:2006fk} to remove sequences with more than 80\% of sequence similarity. This operation resulted in a data set of $45$ sequences.

We design our benchmark using a leave-one-out strategy. We randomly picked one sequence from our data set and performed $12$ random
mutations. Our sequences have $119$ nucleotides, thus the number of mutations corresponds to an error-rate of 10\%. We repeated this operation 
$10$ times. The value of $\beta$ has been set to $1.5$ (larger values gave similar results). To estimate the impact of the distribution of the weights
between the energy term and isostericity  score, we use 4 different values of $\alpha = {0, 0.5, 0.8, 1.0}$. Similarly, we also investigate the impact of 
an under- and over- estimate of the number of errors, and use values equal to 50\% (6 mutations) and 200\% (24 mutations) of the exact number of
errors (i.e. 12).

To evaluate our method, we computed a ROC curve representing the performance of a classifier based on the mutational probabilities computed
by \RNApyro. More specifically, we fix a threshold $\lambda \in [0,1]$ and we predict an error at position $i$ in sequence $\omega$ if and only if the
probability $P(i,n)$ of a nucleotide $n \in \{ A,C,G,U \}$ exceed this threshold. To correct the errors we use the set of nucleotides with a probability
higher than this threshold is  $C(i) = \{ n \; | \;  n \in \{ A,C,G,U \} \mbox{ and } P(i,n) > \lambda \mbox{ and }  n \neq \omega[i] \}$, where $\omega[i]$ is
the nucleotide at position $i$ in the input sequence. We note that for the lowest thresholds multiple nucleotides can be available in $C(i)$ to correct
the sequence. Here, we remind that we aim to estimate the potential of error-correction of \RNApyro and not to develop a error-correction pipe-line. 
The latter will be investigate in further studies. Finally, we progressively vary $\lambda$ between $0$ and $1$ to calculate the ROC curve and the area
under the curve (AUC). We report our results in Figure \ref{fig:ROCall}.

>>>>

\begin{table}[b]
\begin{center}
\begin{tabular}{lccc}
$|s|$&\multicolumn{3}{c}{Number of mutations}\\\cline{2-4}
		 			  & 6   &  12  & 24\\\cline{2-4}
100  				& 35s  & 238s & 1023s\\
300  			& 135s & 594s &2460s\\\cline{2-4}
		 						& 25   &    &	50		\\\cline{2-4}
500         & 5400s&       &  21003s    \\
\end{tabular}
\end{center}
\caption{Time to compute all probabilities. The first column indicates the length and  the column indexes indicate the number
 of mutations. $\alpha$ is
set at $0.5$,  $\beta$ to $1.5$ and $|\Omega|=44$.}
\label{tab:time}
\end{table}

@@

We built our data set from the 5S rRNA multiple sequence alignment (MSA) available in the Rfam Database 11.0 (Rfam id: \texttt{RF00001}).
Since our software does not currently implement gaps (mainly because scoring indels is a challenging issue that cannot be fully addressed
in this work),  we clustered together the sequences with identical gap locations. From the $54$ MSAs without gap produced, we selected the
biggest MSA  which contains $130$ sequences (out of $712$ in the original Rfam MSA). Then, in order to avoid overfitting, we used \texttt{cd-hit}
\cite{Li:2006fk} to remove sequences with more than 80\% of sequence similarity. This operation resulted in a data set of $45$ sequences.

We designed our benchmark using a leave-one-out strategy. We randomly picked one sequence from our data set and performed $12$ random
mutations. Our sequences have $119$ nucleotides, thus the number of mutations corresponds to an error-rate of 10\%. We repeated this operation 
$10$ times. The value of $\beta$ has been set to $1.5$ (larger values gave similar results). To estimate the impact of the distribution of the weights
between the energy term and isostericity  score, we used 4 different values of $\alpha = {0, 0.5, 0.8, 1.0}$. Similarly, we also investigated the impact of 
an under- and over- estimate of the number of errors, and use values equal to 50\% (6 mutations) and 200\% (24 mutations) of the exact number of
errors (i.e. 12).

To evaluate our method, we computed a ROC curve representing the performance of a classifier based on the mutational probabilities computed
by \RNApyro. More specifically, we fixed a threshold $\lambda \in [0,1]$, and predicted an error at position $i$ in sequence $\omega$ if and only if the
probability $P(i,n)$ of a nucleotide $n \in \{ A,C,G,U \}$ exceeds this threshold. To correct the errors we used the set of nucleotides with a probability
higher than this threshold is  $C(i) = \{ n \; | \;  n \in \{ A,C,G,U \} \mbox{ and } P(i,n) > \lambda \mbox{ and }  n \neq \omega[i] \}$, where $\omega[i]$ is
the nucleotide at position $i$ in the input sequence. We note that for the lowest thresholds multiple nucleotides can be available in $C(i)$ to correct
the sequence. Here, we remind that our aim is to estimate the potential of error-correction of \RNApyro and not to develop an error-correction pipe-line, which  
will be the subject of further studies. Finally, we progressively varied $\lambda$ between $0$ and $1$ to calculate the ROC curve and the area
under the curve (AUC). We report our results in Figure~\ref{fig:ROCall}.
